Sound Synthesis

### 2021-03-14 THESIS.tex
joined thesis in one tex file

\input{00_Include.tex}

@@

%\include{01_Resumo.tex}
\pagestyle{empty}
\include{02_Abstract.tex}

@@

\newpage
\pagestyle{fancy}
\include{03_Introduction.tex}
\include{04_Theory.tex}
\include{05_Characterization.tex}
\include{06_SignalEnvelope.tex}

>>>>

\documentclass[12pt]{article}

% \usepackage[utf8]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{authblk}

\usepackage[backend=biber]{biblatex}
%\usepackage{biblatex}

\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
\usepackage{indentfirst}
\setlength{\parindent}{1.25cm}

\usepackage[no-math]{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont{Tex Gyre Termes}

\usepackage[nodisplayskipstretch]{setspace}
\setstretch{1.427465}

\usepackage[explicit]{titlesec}
\titleformat{\section}[hang]{\normalfont\bfseries}{\thesection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsection}[hang]{\normalfont}{\thesubsection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsubsection}[hang]{\normalfont\bfseries}{\thesubsubsection}{1em}{#1}%{\MakeLowercase{#1}}

\usepackage{titletoc}
\titlecontents{section}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
\titlecontents{subsection}[1em]{\vspace{12pt}}{\normalfont\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
\titlecontents{subsubsection}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}}{\hspace*{-2em}}{\titlerule*[.75em]{.}\contentspage}

\usepackage{amsmath}  % enables declare math operator
\DeclareMathOperator{\sign}{sign}

\usepackage{amsfonts} % \mathbb
\usepackage{graphicx} % manipulate imported images

\usepackage{fancyhdr} % \pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\usepackage[toc, section=section]{glossaries}
\makeglossaries

\usepackage{hyperref}
\usepackage{hypcap}
\hypersetup{
  colorlinks=true, 
  allcolors=black,
  pdfauthor={Carlos Henrique Tarjano Santos},
  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
}

\bibliography{00_BIBLI}

\title{Sound Synthesis}
\author{Carlos Tarjano}

\newacronym{FPS}{FPS}{frames per second}
\newacronym{LMS}{LMS}{Least Mean Squares}

% \newglossaryentry{VST}{name=VST, description={Virtual Studio Technology}}

@@

%\include{01_Resumo.tex}
\pagestyle{empty}

\begin{abstract} % 150 words

\end{abstract}

{\bf Keywords:}

@@

\newpage
\pagestyle{fancy}

\newacronym{ENIAC}{ENIAC}{Electronic Numeric Integrator and Calculator}

\section{Introduction}
% Tell the reader the problem you are tackling in this project.

Humans are a creative bunch. Since the beginning of computing, even when mainframes were the only available medium, restricted to organizations with considerable budgets, people found ways and motivation to overcome the technical limitations and develop artistic applications, like video games, in their (and the machine's) spare time.

On a parallel note, also relevant to this thesis, those first machines, like the \gls{ENIAC} in 1946, brought with then the promise of the artificial intelligence as their ultimate goal \parencite{2010DonovanReplay}.

Music, on the other hand, is another staple of humanity's creativity. % TODO first musical instruments
 Is only natural that, as soon as this technology became reasonably stable, musical applications started to emerge.

% State clearly how you aim to deal with this problem. 
% Limit the scope of your study. 
% Sketch out how the thesis is structured to achieve your aim. 
% objective
Ambitious as it may seem, it is the goal of this monograph to motivate a change of approach in domains that, despite their modest intersections, share more attributes than is apparent at a superficial glance: Both machine learning and digital musical acoustics are areas that, despite their current relevance, are fresh out of their respective infancies, and a scrutiny of both promptly reveals the difficulties that arise from this state of affairs. A lack of their own terminology, with frequent borrows from other, more established areas, can be readily cited, as the pronounced prominence of few individual contributions.

Those shortcomings have, however, a bright side to them, in that they still expose those areas to reinterpretation, rendering them somewhat open to the influence of outside ideas.

Digital musical acoustics borrows heavily from the digital signal processing terminology, the last having its roots in the analogic world. Many algorithms are conceived in terms of filters, circuits and other similar legacy constructs, a fact that burdens the conceptual, theoretical framework of the area. Besides the introduction of a noveau envelope extraction method, we find that his formulation in terms of (differential) geometric concepts is, perhaps, as great a contribution as the algorithm per se.

% Perhaps less daring, and owing more to the advanced age of the author than to his research prowess, is the presentation of this monograph in a format more akin to a book; In addition to turning this somewhat dense tome into a more palatable experience.

\section{Bibliographic Approach}

with the advance of technology, some classic problems of bibliographic research were mitigates, while new ones were introduced. a modern problem encontered by the author during the development of this work is the fragmentation of data. In the bygone days of physical research, a paper existed in primarily  physical location.
we adress the meta problem of organizing the bibliography introducing a piece of bibliographic software based on an novel paradigm.

neural networks faced many winters since its introduction. The pattern seems to repeat itself: hardware improvements enable attractive new results, the interest rises, and some technical advancements are made. 
Although, in its current renaissance, the field is still blooming, there are some that are already pointing to a decrease in the research output.

We feel that this tool still offers a great potential in its current state of development; a potential thtat will only be fully explored once it is realised that workflows and auxiliary theories are needed, to be used in conjunctions with neural networks.

In this sense, we introduce a novel representation for pseudo periodic time series, of which sound representation are our primary interest. Although much of the theory is drawn from conventional areas related to digital signal processing, image processing, etc, the representation was formulated from the beginning having in mind the strengths and limitations of neural networks in general.

writing a phd thesis is a peculiar endeavour in the academic world, in the sense that it is a requirement that must be fullfilled only once: while a thriving academic carreer pressuposes a constant output of papers, and possibly books, or book chapters, the closest contact an academic has with the format of a thesis is probably during the development of his masters dissertation; that one endeavour being, all in all, substantially different.

It comes as no surprise, then, that bibliographic software, being in general geared towards the task of aiding in the productions of academic articles, can be subpar for a thesis, where the hability to handle a mutch larger volume of bibliographic material is needed.

\section{Conventions and definitions}
The field explored in this work, besides being extensive and fragmented, is largely interdisciplinar, with terminology strong influenced by widespread usage in the community rather than academic tradition. To maintain coherence thougout the text, it is convenient to stablish some definitions, as some of the terms here applyed may deviate from their usage in particular fields.

Effort was made to comply with the most commom nomenclature when terms are encountered frequently in literature.

\section{Theory}

\subsection{Waves}

It can be insightful to analyse the behaviour of the crytical points and the zeroes of an arbitrary, periodic wave over time.
We will assume, for the sake of simplicity, that each cycle of said wave has a unique maximum and minimum, and crosses the horizontal axis one single time. Those assumption dont compromise the generality of the analysis, as in the case of multiple maxima, minima and/or zeroes, one of thosee can be choosen randomly to represent the global one, as long as consistency between cycles is maintained.

We are going to also assume that a wave with a zero phase has an maximum instantaneous amplitude at time zero, as is the case of the cosine wave, for example. Altough that assumption somewhat departs from the physical intuition, it is useful in making our description consistent with the theory behind the Fourier Transform.

For a continuous wave, if we were to plot the time at which any of those values ocurred as a function of their ordinality, the result would be parallel straight lines, their slope somehow related to the frequency of the underlying wave. We will illustrated the point in the case of discrete waves, as this case will be of more immediate use in this work.

\subsubsection{Continuous and discrete waves}

First, it is necessary to establish a correspondence between a continuous and a discrete wave, as the latter will be the object of most of this work. To do that we will define a simple continuous sinusoidal wave of the form $ S(t) = a \cos \left( \phi + 2 \pi \mathcal{F} t \right) $, where $ a \in \mathbb{R}^+ $ is the amplitude; $ \phi \in \mathbb{R}^+, 0 \le \phi < 2 \pi $ is the phase; $ \mathcal{F} \in \mathbb{R}^+ $ is the absolute frequency of the wave, in Hz and $ t \in \mathbb{R} $ is time, in seconds.

The cosine function is used for a twofold reason: its use helps to underline the periodic nature of the wave in consideration and is the parameter of the phase when the FT is used. It is important to keep in mind, however, that this discussion applies any arbitrary function of the form $ f(x) = f(x + P) \forall x \in D(f) $, where $ D $ is the domain of the function $ f $.
% TODO This will be further illustrated in the discrete part below.

If we were to sample $ S(t) $ at regular intervals, that is, to measure the value of the function at a constant rate of $ h $ \gls{FPS}, neglecting errors of measurement, we would have a discrete representation of the continuous wave of the form $ W[i] = a \cos \left( \phi + 2 \pi f \frac{i}{n} \right) $ where, alongside the quantities introduced before, we would have $ i \in \mathbb{N} $ as the frame number; $ n \in \mathbb{N} $ as the number of discrete samples and $ f \in \mathbb{R}^+ $ being the relative frequency of our discrete wave in cycles per $ n $ samples.

The relative, local frequency $ f $ of the discrete representation is related to the absolute frequency $ \mathcal{F} $ via the equation $ f = \frac{n}{\text{fps}} \mathcal{F} $ while the absolute time, in the continuous case can be obtained from the frame number $ i $ of the discrete signal via $ i = t \ \text{fps} $.

\subsubsection{Points of interest of periodic waves}

\subsection{Least Mean Squares}

The linear \gls{LMS}, also known as linear regression \parencite{2018StephenBoydIntroduction} is a data fitting technique. % TODO

We present here the theory for real numbers, noting that it can be easily extended for complex numbers \parencite{1995LawsonSolving}, should the necessity arise. That, however, will not be the case in this work.

Given two vectors $ X $ and $ Y $, such that:
$$ X = [x_0, x_1, \cdots ,x_{m-1}]^T, \ x_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
$$ Y = [y_0, y_1, \cdots ,y_{m-1}]^T, \ y_i \ \in \ \mathbb{R} \ \forall \ i, \ 0 \le i < m \in \mathbb{N} $$
We ideally want to find the vector of coefficients $ A = [a_0, a_1, \cdots ,a_k]^T, \ a_j \ \in \ \mathbb{R} \ \forall \ j, \ 0 \le j \le k \in \mathbb{N} $ that satisfies the equality $ x^0_i a_0 + x^1_i a_1 + \cdots + x^{k-1}_i a_{k-1} + x^k_i a_k = y_i \quad \forall \quad i \in \{0, \cdots , m-1\} $.

Put in that way, what we have is a system of linear equations that can be rewritten in matrix form as seen in equation \ref{eq:lms}.

\begin{equation}
  \label{eq:lms}
\stackrel{Q_{m \times (k+1)}}
{
\begin{bmatrix}
1 & x_0 & \cdots & x_0^k \\
1 & x_1 & \cdots & x_1^k \\
\vdots & \vdots & \ddots & \vdots  \\
1 & x_{m-1} & \cdots & x_{m-1} \\
\end{bmatrix}
}
\stackrel{A_{(k+1) \times 1}}
{
\begin{bmatrix}
a_0 \\
a_1 \\
\vdots  \\
a_k \\
\end{bmatrix}
}
=
\stackrel{Y_{m \times 1}}
{
\begin{bmatrix}
y_0 \\
y_1 \\
\vdots  \\
y_{m-1} \\
\end{bmatrix}
}
\end{equation}

More compactly, we have that $ Q A = Y $. This system will generally be over-determined, only having an exact solution in the case where $ Y $ is a linear combination of the columns in $ Q $ \parencite{2018StephenBoydIntroduction}.

Since an exact solution is often beyond reach, we can at least choose $ A $ in a way that makes $ Q A $ and $ Y $ close. A way to do that is to minimize the absolute errors (or their square) between those two quantities; $ \underset{A}{\min} ||Q A - Y||^2_2 $

Using calculus, for example, on can arrive at the equation $ \hat{A} = (Q^T Q)^{-1} Q^T Y $, where $ \hat{A} $ is the vector of coefficients that minimizes the squared errors. A detailed derivation is available at \cite{2018StephenBoydIntroduction}.

\subsection{Constrained Least Mean Squares}

\newacronym{CLMS}{CLMS}{Constrained Least Mean Squares}

In the constrained version of the \gls{LMS}, known as the \gls{CLMS} problem we are still interested in finding the coefficient vector $ \hat{A} $ that minimizes $ ||Q A - Y||^2_2 $, but respecting a restriction on this vector of the general form of $ V A = B $.

We can then define the Lagrangian and set its derivatives to zero to obtain, as described in \textcite{2013SelesnickLeast}, the closed form solution

$$ \hat{A} = \left(Q^{T} Q\right)^{-1} \left(Q^{T} Y - V^{T} \left(V \left(Q^{T} Q\right)^{-1} V^{T}\right)^{-1} V \left(Q^{T} Q\right)^{-1} Q^{T} Y - B \right) $$

\subsection{Savitzky–Golay filter}

\section{Characterization of a digital Signal}
% TODO continuous to digital

\subsection{Envelope of a digital signal}

Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a smooth function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. Nevertheless, a general method to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge about the nature of the specific waves under investigation.
To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, completely eliminating the necessity of parameter tuning.
The approach here described draws inspiration from geometric concepts to isolate the frontiers and thus estimate the temporal envelope of an arbitrary signal; to that end, alpha-shapes, concave hulls, and discrete curvature are explored. We also define entities, such as a pulse and frontiers, in the context of an arbitrary digital signal, as a means to reduce dimensionality and the complexity of the proposed algorithm. Specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, serving as a threshold to identify the wave's frontier points. 
We find the algorithm accurate in the identification of the frontiers of a wide range of digital sound waves with very diverse characteristics, while localizing each pseudo-cycle of the wave in the time domain. The algorithm also compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform.
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.

\subsubsection{Envelope Detection}
Envelope detection is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

In many natural signals, however, the temporal amplitude envelope of a signal plays a prominent role in the characteristics exhibited: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of mandarin tones. This is also the case for English \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.

The envelope helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.

When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

Addressing the different units in the horizontal and vertical axes, one can transform the DSP problem of envelope detection into the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $. For that purpose, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, a mathematically well-defined extension to the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

This approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of cross-fertilization.

Steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the wave, and also via the direct translation of computer vision methods to the task \parencite{2015YangRepresenting}.

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid the need for manual intervention or parameter tuning.

While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.

The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis. Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.

% The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.

We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.

We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.

\subsection{Frequency content}

\subsection{Representations}

% one hot

\section{Envelope Detection}

Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
% Problem | Objectives
Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope. 
% Results | Findings | Product
The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
% Conclusions | Implications
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. We briefly discuss some potential paths in this direction. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.

Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones. This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.

When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $. Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[H]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.

Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.

The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.

\subsection{Methods}
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.

\subsubsection{Characterization of a temporal envelope}

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude}.

Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).

In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.

We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$. This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.

The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second. We can thus define:

\begin{subequations}\label{eq:Envelope}
\begin{align}
  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\  
  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{align}
\end{subequations}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$. Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.

Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.

This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave. Only the extrema touched by the circle will be marked as part of the envelope.

In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

\begin{figure}[H]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

\subsubsection{Simplified Representation of a Wave}

Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.

This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.

\begin{figure}[H]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.

To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points. We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.

The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:
